\herschel\thanks{{\it Herschel} is an ESA space observatory with science instruments provided by European-led Principal Investigator consortia and with important participation from NASA.} Orion Protostar Survey (HOPS) Imaging and Photometry Results.

### 2017-04-15 HOPS/photometry/ms.tex
rev 4 on the way

\subsection{Flux ratio distribution}
\label{sec:fratio}
\par
Flux ratios (colors) are better suited for determining differences as a function of source type than the flux itself because flux ratios depend on the shape of the source SED \citep{ali}. We investigated two different flux ratios using the combined Spitzer, HOPS and Gould-Belt photometry: \clra, and \clrb. Figures~\ref{fig:clrclr1}~\&~\ref{fig:clrclr2} show the flux-ratio vs flux-ratio plots segregated by sub-region for Orion A and Orion B. We do not include data with upper limits for these plots (sources with flag value = 2 in Table~\ref{tbl:obs}). The red symbols show the colors for the sub-region identified in the sub-panel, and the black symbols show the colors for the entire sample. As with the median of the FD, we note an apparent systematic change in the observed flux ratios relative to the entire population in both Orion~A and Orion~B.
\par
We quantitatively explore this apparent trend further in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec}. These Figures show the mean and median colors for each of the sub-regions as a function of the mean declination for all sources belonging to the respective sub-region.  Since the sub-regions are vertically segregated in both the Orion~A and B clouds, declination is a useful proxy for sub-region.
\par
In Orion~A, the median and mean colors of the sub-regions become systematically redder towards higher declinations.  In Orion~B, the median and mean colors of the sub-regions becomes systematically redder towards lower declinations.  For both, the vertical bar shows the 1-$\sigma$\ standard deviation of the flux-ratio values as a proxy for the range.  Even with large variation in color, the systematic trends with declination is clearly delineated.

@@

Completeness limits are difficult to estimate accurately.  The detection efficiency is a strong function of the magnitude of the local nebular emission;  Hence, we expect the limits to vary considerably even within a single sub-region as the nebular emission itself varies inside any of our adopted sub-region boundaries.  This variation makes it difficult to disentangle observational bias from any real trends shown in Figures~\ref{fig:mdFlux_v_dec},~\ref{fig:clrclr1},~\&~\ref{fig:clrclr2}.

@@

Fortunately, there is a way to decouple detection bias.   Differences in completeness limit will affect the median flux distribution (Figure~\ref{fig:mdFlux_v_dec}), since such differences will bias the median calculation.  For the same effect to exist also in the flux ratio correlation (Figures~\ref{fig:clrclr1}~\&~\ref{fig:clrclr2}) requires that the observed flux be correlated with flux ratio.  We investigate this possibility by directly examining  the data for such a correlation, and via a simple simulation described below.

@@

We can now further investigate completeness limit and observational bias via a simple simulation.  The LDN 1641 sub-regions contain the largest number of protostars (see Table~\ref{tbl:regions}) and visual inspection shows the least contamination of the sample from extended nebulosity. Further, fainter protostars are detected in LDN 1641 than in Orion OMC 2/3. Hence, if bright, complex, spatially extended emission from the local dust is responsible for decreased reliability and higher completeness limits for point source photometry, then LDN 1641 provides an ideal contrast to the Orion OMC 2/3 sub-region. The contrast between LDN~1641 and Orion OMC-2/3 is at the extreme ends of the spatial correlation observed in Figures~\ref{fig:mdFlux_v_dec}$-$~\ref{fig:clrclr2}.

@@

\begin{figure*}[ht]
\centering
\includegraphics[height=7in]{figures/LF70_byregion.eps}
\caption{The 70\micron\ flux distributions segregated by sub-region for Orion B (left column) and Orion A (right column).\label{fig:fd70regional}}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\includegraphics[height=7in]{figures/LF160_byregion.eps}
\caption{The 160\micron\ flux distributions segregated by subregion for Orion B (left column) and Orion A (right column).\label{fig:fd160regional}}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\plotone{figures/OrionA_clrclr.eps}
\caption{The observed distribution of colors in Orion~A segregated by sub-regions.  The sub-region is identified in the respective panel.  Solid red symbols show the colors for the sub-region.  Smaller black open symbols show the distribution for all sources in the combine Orion~A and Orion~B samples.\label{fig:clrclr1}}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\plotone{figures/OrionB_clrclr.eps}
\caption{Same as Figure~\ref{fig:clrclr1} for Orion~B.\label{fig:clrclr2}}
\end{figure*}

>>>>

\subsection{Flux ratio distribution}
\label{sec:fratio}
\par
Flux ratios (colors) are better suited for determining differences as a function of source type than the flux itself because flux ratios depend on the shape of the source SED \citep{ali}. We investigated two different flux ratios using the combined Spitzer, HOPS and Gould-Belt photometry: \clra, and \clrb. Figure~\ref{fig:clrclr} shows the flux-ratio vs flux-ratio plots segregated by sub-region for Orion A and Orion B. We do not include data with upper limits for these plots (sources with flag value = 2 in Table~\ref{tbl:obs}). The red symbols show the colors for the sub-region identified in the sub-panel, and the black symbols show the colors for the entire sample. As with the median of the FD, we note an apparent systematic change in the observed flux ratios relative to the entire population in both Orion~A and Orion~B.
\par
We quantitatively explore this apparent trend further in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec}. These Figures show the mean and median colors for each of the sub-regions as a function of the mean declination for all sources belonging to the respective sub-region.  Since the sub-regions are vertically segregated in both the Orion~A and B clouds, declination is a useful proxy for sub-region.
\par
In Orion~A, the median and mean colors of the sub-regions become systematically redder towards higher declinations.  In Orion~B, the median and mean colors of the sub-regions becomes systematically redder towards lower declinations.  For both, the vertical bar shows the 1-$\sigma$\ standard deviation of the flux-ratio values as a proxy for the range.  Even with large variation in color, the systematic trends with declination is clearly delineated.

@@

Completeness limits are difficult to estimate accurately.  The detection efficiency is a strong function of the magnitude of the local nebular emission;  Hence, we expect the limits to vary considerably even within a single sub-region as the nebular emission itself varies inside any of our adopted sub-region boundaries.  This variation makes it difficult to disentangle observational bias from any real trends shown in Figures~\ref{fig:mdFlux_v_dec},~\&~\ref{fig:clrclr}.

@@

Fortunately, there is a way to decouple detection bias.   Differences in completeness limit will affect the median flux distribution (Figure~\ref{fig:mdFlux_v_dec}), since such differences will bias the median calculation.  For the same effect to exist also in the flux ratio correlation (Figure~\ref{fig:clrclr}) requires that the observed flux be correlated with flux ratio.  We investigate this possibility by directly examining  the data for such a correlation, and via a simple simulation described below.

@@

We can now further investigate completeness limit and observational bias via a simple simulation.  The LDN 1641 sub-regions contain the largest number of protostars (see Table~\ref{tbl:regions}) and visual inspection shows the least contamination of the sample from extended nebulosity. Further, fainter protostars are detected in LDN 1641 than in Orion OMC 2/3. Hence, if bright, complex, spatially extended emission from the local dust is responsible for decreased reliability and higher completeness limits for point source photometry, then LDN 1641 provides an ideal contrast to the Orion OMC 2/3 sub-region. The contrast between LDN~1641 and Orion OMC-2/3 is at the extreme ends of the spatial correlation observed in Figures~\ref{fig:mdFlux_v_dec}$-$~\ref{fig:clrclr}.

@@

\begin{figure*}[ht]
\centering
\includegraphics[height=7in]{figures/LF70_byregion.eps}
\caption{The 70~\micron\ flux distributions segregated by sub-region for Orion B (left column) and Orion A (right column).\label{fig:fd70regional}}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\includegraphics[height=7in]{figures/LF160_byregion.eps}
\caption{The 160~\micron\ flux distributions segregated by subregion for Orion B (left column) and Orion A (right column).\label{fig:fd160regional}}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\plotone{figures/clrclr.eps}
\caption{The observed distribution of colors in Orion~A (right panel) and Orion~B (left panel) segregated by sub-regions.  The sub-region is identified in each respective panel.  Solid red symbols show the colors for the sub-region.  Smaller black open symbols show the distribution for all sources in the combined Orion~A and Orion~B samples. \label{fig:clrclr}}
\end{figure*}

@@

%\begin{figure*}[ht]
%\centering
%\plotone{figures/OrionA_clrclr.eps}
%\caption{The observed distribution of colors in Orion~A segregated by sub-regions.  The sub-region is identified in the respective panel.  Solid red symbols show the colors for the sub-region.  Smaller black open symbols show the distribution for all sources in the combine Orion~A and Orion~B samples.\label{fig:clrclr1}}
%\end{figure*}
%
%\clearpage
%
%\begin{figure*}[ht]
%\centering
%\plotone{figures/OrionB_clrclr.eps}
%\caption{Same as Figure~\ref{fig:clrclr1} for Orion~B.\label{fig:clrclr2}}
%\end{figure*}
%
%\clearpage
